---
title: "Getting Started with fdrregGenomics"
author: "Jinghong Qiu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with fdrregGenomics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

## Overview

**fdrregGenomics** provides a comprehensive pipeline for integrating GWAS
summary statistics with biological annotations using False Discovery Rate
Regression (FDRreg). It supports three genomic levels:

| Tier | Function | Description |
|------|----------|-------------|
| SNP | `run_fdrreg_snp()` | SNP-level GWAS with LDSC decorrelation |
| MAGMA Gene | `run_fdrreg_magma_gene()` | MAGMA gene-level results |
| TWAS | `run_fdrreg_twas()` | S-PrediXcan / S-MultiXcan results |

### Disclaimer / Attribution

> **The underlying FDRreg algorithm is developed by Scott et al. (2016);
> this package provides the comprehensive genomics integration,
> preprocessing, and validation workflow.**

### References

Scott, J. G., Kelly, R. C., Smith, M. A., Zhou, P., & Kass, R. E. (2016).
False discovery rate regression: an application to neural synchrony
detection in primary visual cortex.
*Journal of the American Statistical Association*, 110(510), 459-471.

---

## Installation

```{r install, eval = FALSE}
# Install from local source or GitHub
# install.packages("fdrregGenomics")
```

## Simulated Example Data

We use `simulate_example_data()` to create realistic simulated datasets.

```{r simulate}
library(fdrregGenomics)

sim <- simulate_example_data(
  n_snps   = 2000,
  n_genes  = 200,
  n_aux    = 3,
  n_annot  = 10,
  prop_signal = 0.03,
  seed = 42
)

cat("Target GWAS SNPs:", nrow(sim$snp_target), "\n")
cat("Auxiliary traits:", names(sim$snp_aux), "\n")
cat("Overlapping traits:", sim$overlap_traits, "\n")
cat("Annotations:", ncol(sim$annotations) - 1, "columns\n")
```

## Advanced Simulation Examples

### Multiple Simulation Modes

```{r sim_modes}
library(fdrregGenomics)

# Mode 1: Full simulation (default)
sim_full <- simulate_example_data(
n_snps = 1000,
simulation_mode = "full",
seed = 42
)

# Mode 2: Summary statistics only
sim_summary <- simulate_example_data(
n_snps = 1000,
simulation_mode = "summary_only",
signal_model = "complex",
# Make sure function arguments match n_annot
signal_function = function(x1, x2, x3) {
 -3.0 + 0.8*x1 + 1.0*x2 + 1.2*x3
},
n_annot = 3, # Match number of function arguments
seed = 42
)

# Mode 3: Raw data only
sim_raw <- simulate_example_data(
n_snps = 1000,
simulation_mode = "raw_only",
seed = 42
)

cat("Full mode output length:", length(sim_full), "\n")
cat("Summary mode SNP data rows:", nrow(sim_summary$snp), "\n")
cat("Raw mode feature matrix dim:", dim(sim_raw$features), "\n")
```

### Complex Signal Models

```{r complex_signal}
# Define a complex signal model with 5 parameters
complex_model <- function(x1, x2, x3, x4, x5) {
-2.5 + 0.5*x1 + 0.8*x2 + 1.2*x3 - 0.3*x4 + 0.1*x5
}

sim_complex <- simulate_example_data(
n_snps = 2000,
simulation_mode = "summary_only",
signal_model = "complex",
signal_function = complex_model,
n_annot = 5, # Match number of function arguments
seed = 42
)

# Check signal rate
signal_rate <- mean(sim_complex$true_info$snp$is_signal)
cat("Complex model signal rate:", round(signal_rate, 3), "\n")
```
```

---

## Example 1: SNP-Level Analysis

### Basic SNP Pipeline (with Sample-Overlap Decorrelation)

```{r snp_basic}
result_snp <- run_fdrreg_snp(
  target          = sim$snp_target,
  aux             = sim$snp_aux,
  overlap_traits  = sim$overlap_traits,
  ldsc_intercepts = sim$ldsc_intercepts,
  id_col          = "snpid",
  z_col           = "z",
  p_col           = "pval",
  feature_transform = "signed",
  var_select      = "none",
  fdrreg_nulltype = "both",
  seed            = 42
)

# Print the result
print(result_snp)
```

### Access Significant Findings

```{r snp_sig}
sig_theo <- significant(result_snp, threshold = 0.05, type = "theoretical")
cat("Significant SNPs (theoretical, FDR < 0.05):", nrow(sig_theo), "\n")

head(sig_theo)
```

### Full Results Table

```{r snp_full}
full <- result_snp$full_results
cat("Columns:", paste(colnames(full), collapse = ", "), "\n")
cat("Dimensions:", nrow(full), "x", ncol(full), "\n")
```

### With LASSO Variable Selection

```{r snp_lasso}
result_lasso <- run_fdrreg_snp(
  target          = sim$snp_target,
  aux             = sim$snp_aux,
  overlap_traits  = sim$overlap_traits,
  ldsc_intercepts = sim$ldsc_intercepts,
  annotations     = sim$annotations,
  feature_transform = "signed",
  var_select      = "lasso",
  seed            = 42
)

print(result_lasso)
```

### With Biological Annotations Only

```{r snp_annot}
result_ann <- run_fdrreg_snp(
  target      = sim$snp_target,
  annotations = sim$annotations,
  feature_transform = "signed",
  var_select  = "none",
  seed        = 42
)

print(result_ann)
```

### Using Pre-Built Covariance Matrix

```{r snp_covmat}
result_cov <- run_fdrreg_snp(
  target         = sim$snp_target,
  aux            = sim$snp_aux,
  overlap_traits = sim$overlap_traits,
  cov_matrix     = sim$cov_matrix,   # instead of ldsc_intercepts
  feature_transform = "signed",
  seed           = 42
)

summary(result_cov)
```

---

## Example 2: MAGMA Gene-Level Analysis

```{r magma}
result_magma <- run_fdrreg_magma_gene(
  target            = sim$magma_target,
  aux               = sim$magma_aux,
  annotations       = sim$gene_annotations,
  id_col            = "GENE",
  z_col             = "ZSTAT",
  p_col             = "P",
  feature_transform = "abs",
  var_select        = "lasso",
  fdrreg_nulltype   = "both",
  seed              = 42
)

print(result_magma)

# Significant genes
sig_genes <- significant(result_magma, threshold = 0.05)
cat("Significant genes:", nrow(sig_genes), "\n")
```

---

## Example 3: TWAS Analysis

### S-PrediXcan

```{r twas_predi}
result_predi <- run_fdrreg_twas(
  target          = sim$spredixcan_target,
  twas_type       = "spredixcan",
  aux             = sim$spredixcan_aux,
  annotations     = sim$gene_annotations,
  feature_transform = "abs",
  var_select      = "elasticnet",
  seed            = 42
)

print(result_predi)
```

### S-MultiXcan

```{r twas_multi}
result_multi <- run_fdrreg_twas(
  target          = sim$smultixcan_target,
  twas_type       = "smultixcan",
  aux             = sim$smultixcan_aux,
  annotations     = sim$gene_annotations,
  feature_transform = "abs",
  var_select      = "lasso",
  seed            = 42
)

print(result_multi)
```

---

## Example 4: Unified Dispatcher

```{r unified}
# The run_fdrreg() function dispatches to the appropriate wrapper
result <- run_fdrreg(
  tier = "twas",
  target    = sim$spredixcan_target,
  twas_type = "spredixcan",
  aux       = sim$spredixcan_aux,
  seed      = 42
)

summary(result)
```

---

## Comparing Variable Selection Methods

```{r comparison, eval = FALSE}
methods <- c("none", "lasso", "marginal", "elasticnet")

results_list <- lapply(methods, function(m) {
  run_fdrreg_snp(
    target          = sim$snp_target,
    aux             = sim$snp_aux,
    overlap_traits  = sim$overlap_traits,
    ldsc_intercepts = sim$ldsc_intercepts,
    feature_transform = "signed",
    var_select      = m,
    seed            = 42
  )
})
names(results_list) <- methods

# Compare discoveries at FDR < 0.05
sapply(results_list, function(r) {
  sum(r$full_results$fdr_theo < 0.05, na.rm = TRUE)
})
```

---

## Understanding the Output Object

The `fdrreg_result` S3 object contains:

| Slot | Description |
|------|-------------|
| `$tier` | Analysis tier (snp, magma_gene, twas_*) |
| `$full_results` | Data frame: id, z, p, fdr_theo, lfdr_theo, fdr_emp, lfdr_emp, pep_* |
| `$summary` | Data frame: discovery counts at standard FDR thresholds |
| `$assessment` | Data frame: feature beta, SE, z-score, p-value from model |
| `$models` | List of fitted FDRreg model objects |
| `$varselect_model` | Variable selection model (or NULL) |
| `$covariates_used` | Character vector of feature names used |
| `$params` | Reproducibility: seed, R version, package versions, timestamp |

```{r output_demo}
names(result_snp)
result_snp$summary
result_snp$params
```
